\ProjectName{}: Data Imputation as a Query Optimization

### 2017-02-16 paper/main.tex
paper: \verb to \lstinline, add syntax highlighting
paper: moved motivating example before related work. Cleaned up and added one todo
paper: rewrote abstract from class project to vldb version

\documentclass{vldb}
\usepackage{balance}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage[activate={true,nocompatibility},final,kerning=true,spacing=true,factor=1100,stretch=10,shrink=10]{microtype}
\usepackage[subtle]{savetrees}
\usepackage{amsfonts,amssymb,amsmath}
\usepackage{xcolor}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage[style=numeric,firstinits=true]{biblatex}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{hyperref} % load after other packages
\usepackage{graphicx} % includegraphics
\usepackage{listings} % code samples with syntax hl
\usepackage{qtree} % for simple tree diagrams
\usepackage{subcaption}
\addbibresource{references.bib}

@@

% listings settings
\lstset{ %
  backgroundcolor=\color{white},   % choose the background color; you must add \usepackage{color} or \usepackage{xcolor}; should come as last argument
  basicstyle=\ttfamily\footnotesize,        % the size of the fonts that are used for the code
  breakatwhitespace=false,         % sets if automatic breaks should only happen at whitespace
  breaklines=true,                 % sets automatic line breaking
  captionpos=b,                    % sets the caption-position to bottom
  commentstyle=\color{red},    % comment style
  %frame=single,	                   % adds a frame around the code
  keepspaces=true,                 % keeps spaces in text, useful for keeping indentation of code (possibly needs columns=flexible)
  keywordstyle=\color{blue},       % keyword style
%  numbers=left,                    % where to put the line-numbers; possible values are (none, left, right)
  numbersep=5pt,                   % how far the line-numbers are from the code
%  numberstyle=\tiny\color{gray}, % the style that is used for the line-numbers
  rulecolor=\color{black},         % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. comments (green here))
%  stepnumber=2,                    % the step between two line-numbers. If it's 1, each line will be numbered
  stringstyle=\color{mauve},     % string literal style
  tabsize=2,	                   % sets default tabsize to 2 spaces
}

@@

\begin{abstract}
In order to study the placement of an imputation step, we create a logical imputation operator (along with respective physical instances) and incorporate its placement as part of the query plan optimization process into a simple database prototype.  We introduce measures for information loss and runtime for imputation operations, which outline the main trade-offs in the imputation placement. We add these measures into our cost estimation, allowing us to intelligently place the data imputation step during query planning. We show the trade-offs between efficiency and accuracy for simple data imputation models. \end{abstract}

\input{introduction}
\input{relatedwork}
\input{examples}
\input{algorithm}
\input{experiments}
\input{conclusion}

>>>>

\documentclass{vldb}
\usepackage{balance}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage[activate={true,nocompatibility},final,kerning=true,spacing=true,factor=1100,stretch=10,shrink=10]{microtype}
\usepackage[subtle]{savetrees}
\usepackage{amsfonts,amssymb,amsmath}
\usepackage{xcolor}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage[style=numeric,firstinits=true]{biblatex}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{hyperref} % load after other packages
\usepackage{graphicx} % includegraphics
\usepackage{listings} % code samples with syntax hl
\usepackage{stmaryrd}
\usepackage{qtree} % for simple tree diagrams
\usepackage{subcaption}
\addbibresource{references.bib}

@@

% listings settings
\lstset{ %
  backgroundcolor=\color{white},   % choose the background color; you must add \usepackage{color} or \usepackage{xcolor}; should come as last argument
  basicstyle=\ttfamily\footnotesize,        % the size of the fonts that are used for the code
  breakatwhitespace=false,         % sets if automatic breaks should only happen at whitespace
  breaklines=true,                 % sets automatic line breaking
  captionpos=b,                    % sets the caption-position to bottom
  commentstyle=\color{red},    % comment style
  %frame=single,	                   % adds a frame around the code
  keepspaces=true,                 % keeps spaces in text, useful for keeping indentation of code (possibly needs columns=flexible)
  keywordstyle=\color{blue},       % keyword style
%  numbers=left,                    % where to put the line-numbers; possible values are (none, left, right)
  numbersep=5pt,                   % how far the line-numbers are from the code
%  numberstyle=\tiny\color{gray}, % the style that is used for the line-numbers
  rulecolor=\color{black},         % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. comments (green here))
%  stepnumber=2,                    % the step between two line-numbers. If it's 1, each line will be numbered
  stringstyle=\color{mauve},     % string literal style
  tabsize=2,	                   % sets default tabsize to 2 spaces
  language=SQL
}

@@

\begin{abstract}
Missing values abound in data analysis and often present a serious challenge for users, who
are forced to pick between removing tuples with missing values or creating a clean duplicate of their data
by  applying a potentially expensive imputation strategy. We show that it is possible
to incorporate imputation into a cost-based query optimizer, performing imputation on the fly
for each query. This allows users to immediately explore
their data, while the system picks the optimal placement of imputation operations based
on the specific query issued and a parameter that the user selects for evaluating the tradeoff between data
quality and runtime performance. We incorporate the imputation operators into the relational algebra, allowing
for traditional optimizations to take place. Our experiments evaluate this approach on two real-world survey-based datasets
from the \textit{Center for Disease Control and Prevention} and \textit{freeCodeCamp}, showing that the 
optimizer generates query plans that execute in a fraction of the time that it takes to impute values at the base
tables, making exploratory data analysis using our system feasible. Furthermore, we show that the query results
from on-the-fly imputation are comparable to those generated with the standard full-table imputation approach.
\end{abstract}

\input{introduction}
\input{examples}
\input{relatedwork}
\input{algorithm}
\input{experiments}
\input{conclusion}
